\Large\textbf{Preparation GUIDE for Initial-Submission Manuscript}\\~\\ \large\textbf{ Preparation of Papers in Two-Column Format\\ for the ASP-DAC 2014 (\LaTeX2e version)}

### 2013-11-29 delays.tex
Updated text and added discussion text

@@

The aforementioned variants are special cases of the general case
\texttt{delay (M..N)}.

@@

\subsection{The compiler flow}
\label{sec:systemj-tool-chain}

\begin{figure}[t!]
  \centering
  \includegraphics[scale=0.5]{./tool}
  \caption{The tool-chain flow of SystemJ compiler with delays}
  \label{fig:4}
\end{figure}

@@

\subsection{Programming using the delay construct}
\label{sec:progr-using-delay}

In this section we provide a number of examples to show the different
types of real-time programming paradigms that can be incorporated into
the GALS (and its sub-set synchronous) programming model.

\subsubsection{Non-deterministic time}
\label{sec:non-determ-time}

Many real-world systems require non-deterministic timing
constructs. Where the exact real-time delay is not known or should not
be known apriori. One such system is our motivating example -- the human
response time system.

\subsubsection{Timeout}
\label{sec:timeout}

There are many instances when one would like to wait on an input from
environment for only a specified amount of time. This can be programmed
as:
% \begin{scriptsize}
\begin{verbatim}
// timeout after 1ms.
trap(T){{await(A);}||
        {delay(1..1 ms);exit(T);}}
\end{verbatim}
% \end{scriptsize}

\subsubsection{Periodic reactions}
\label{sec:periodic-reactions}

A reaction, or a whole CD, can be programmed to run periodically like
so:

% \begin{scriptsize}
\begin{verbatim}
while(true) { delay ((1..1)ms); emit S; 
 //do something }
\end{verbatim}
% \end{scriptsize}
A periodic reaction (or CD) requires special consideration. Since the
delay statement is converted into \texttt{pause} constructs. One should
not introduce extra \texttt{pause} constructs when building periodic
reactions (or CDs). This is essential since introducing additional
pauses would introduce more logical ticks.

\subsubsection{Interaction of preemption and delays}
\label{sec:inter-preempt-delays}

Preemption plays an important role in reactive languages. One needs to
carefully consider the interplay of \texttt{delay} semantics with the
preemption semantics of reactive languages. Previous attempts at
incorporating delays (using external timers) have only been partially
successful, because of the complex interplay between real-time and
preemption. Consider the simple example below, which models real-time
using external timers as in~\cite{rsh94}.

% \begin{scriptsize}
\begin{verbatim}
suspend(S) {emit START_TIMER(10); 
            await (TIMER);emit O1;}
\end{verbatim}
% \end{scriptsize}

As identified in~\cite{Bourke2009a} \texttt{suspend} does not play well
with the external timer. The above program sends a signal to an external
timer and waits for 10 ms to pass by. Consider what happens when signals
\texttt{S} and \texttt{TIMER} occur in the same logical tick, the
\texttt{await} statement is never executed (due to suspend) and hence,
we enter a deadlock. Such problems are completely avoided in our
technique, because we convert the real-time delays into logical delays
(\texttt{pause} constructs), which interact well with preemption.

\subsubsection{Interaction of channel communication and delays}
\label{sec:inter-chann-comm}

Channels, used for communication between reactions in asynchronously
running CDs, are an addition in the SystemJ language. Like interaction
of preemption and delays, conversion of real-time delays to logical
ticks also interacts well with channel rendezvous, because the semantics
of interaction are well defined~\cite{amal10}. More importantly, we need
to consider the interplay of channel communication and WCRT/BCRT
analysis. Since, channel communication does not stop logical time
(see~\cite{amal10}) WCRT/BCRT are unaffected by channel
communication. % The response time to input signals \textit{is} though!
% But, we are unconcerned with the response time, analysis in this paper
% and it remains a future research avenue.

>>>>

\begin{figure}[t!]
  \centering
  \includegraphics[scale=0.6]{./sem}
  \caption{Pictorial representation of the semantics of the delay
    statement and its variants}
  \label{fig:5}
\end{figure}

@@

Figure~\ref{fig:5} pictorially represents the semantics of these delay
constructs. The construct $delay (M..N)$ requires that once executed the
control-flow suspends execution for \textit{at\_least} $M$ units and
\textit{at\_most} $N$ units, depicted as bounded box in
Figure~\ref{fig:5}. There is no upper bound for the second delay form
$delay(M)$, which requires that the control-flow suspends for
\textit{at\_least} $M$ units, but the following statement can be
executed after some countable delay, depicted with the unbounded box in
Figure~\ref{fig:5}. Finally, $delay(M..M)$, requires that there is an
\textit{exact} delay of $M$ units, i.e., the control-flow suspends
execution for \textit{at\_least} and \textit{at\_most} $M$ units.

@@

\subsection{The tool-chain flow}
\label{sec:tool-chain-flow}

\begin{figure}[h!]
  \centering
  \includegraphics[scale=0.45]{./tool.pdf}
  \label{fig:4}
  \caption{The tool-chain flow of SystemJ compiler with delays}
\end{figure}

@@

Our compiler has a switch that enables over-approximation as described
in Section~\ref{sec:extend-tehcn-vari} if the delay requirements are not
met. If this compiler switch is not enabled the compiler simply outputs
an error indicating the relaxation required. The system designer can
then either set the compiler switch of change the program by shortening
the critical path by introducing \texttt{pause} constructs. We show how
introducing \texttt{pause} constructs can help meet timing requirements
explicitly, using examples, later on in
Section~\ref{sec:experimental-results}.

### 2013-11-29 motivation.tex
Updated paper
Updated text and added discussion text

\begin{verbatim}
{ // Clock-domain 1
while(true) {
  trap(TRAP){
   // Abort if user touches the screen
   abort(TOUCH){
    {sustain GREEN_LIGHT;}
    || // synchronous parallel operator
    {
     //exit after any where from 50.3 to 200.3 ms.
     delay (50.3 .. 200.3 ms);
     exit(TRAP); 
    }
   } do {send S(1);} // Send 1 if touched 
  } do {send S(0);} // Else Send 0
  pause; // only construct to consume time
 }
}
>< // asynchronous operator
{ // Clock-domain 2
  while(true){ receive S; // do something with S}
}
\end{verbatim}
  \end{scriptsize}
\end{minipage}
\caption{An example human responsiveness system (HRTCS)}
% }
% \end{center}
\label{fig:1}
\end{figure}

@@

The SystemJ psuedocode~\cite{amal10} for such a system is shown in
Figure~\ref{fig:1}. We use the SystemJ language in this paper, first of
all, because being a \textit{Globally Asynchronous Locally Synchronous}
(GALS) language it helps us easily describe a larger class of systems
(like HRTCS in Figure~\ref{fig:1}) compared to purely synchronous
languages like Esterel. Moreover, being a super-set of Esterel, SystemJ
allows us to explore the marriage of real-time and logical time, which
is equally applicable to the synchronous sub-set.

>>>>

\begin{verbatim}
{ // Clock-domain 1
while(true) {
  trap(TRAP){
   // Abort if user touches the screen
   abort(TOUCH){
    {sustain GREEN_LIGHT;} //reaction R1
    || // synchronous parallel operator
    {
     //exit after any where from 50.3 to 200.3 ms.
     delay (50.3 .. 200.3 ms); 
     exit(TRAP); 
    } // reaction R2
   } do {send S(1);} // Send 1 if touched 
  } do {send S(0);} // Else Send 0
  pause; // only construct to consume logical time
 }
}
>< // asynchronous operator
{ // Clock-domain 2
  while(true){ receive S; // do something with S}
}
\end{verbatim}
  \end{scriptsize}
\end{minipage}
\caption{An example human responsiveness system (HRTCS)}
% }
% \end{center}
\label{fig:1}
\end{figure}

@@

The SystemJ pseudo-code~\cite{amal10} for such a system is shown in
Figure~\ref{fig:1}. We use the SystemJ language in this paper, first of
all, because being a \textit{Globally Asynchronous Locally Synchronous}
(GALS) language it helps us easily describe a larger class of systems
(like HRTCS in Figure~\ref{fig:1}) compared to purely synchronous
languages like Esterel. Moreover, being a super-set of Esterel, SystemJ
allows us to explore the marriage of real-time and logical time, which
is equally applicable to the synchronous sub-set.
